\bf A Hierarchical Bayesian Approach for Modeling Infant-Mortality and Wearout Failure Modes

### 2017-07-06 paper/paper.tex
Edits to Abstract, Motivation, and Model Section.

\bigskip
\begin{abstract}
Consumers have an interest in accurate assessment of product reliability. Toward this end, there is a need for models that can fit the data together with methods that account for uncertainty. While computationally convenient, simple parametric models can fail to capture key characteristics found in real data.

We illustrate this method on real data with left-truncation and right-censoring where the information in the data is unevenly distributed across the populations of interest.
\end{abstract}

@@

Failure distributions of high-reliability consumer products can be difficult to assess.  Even when field data are available, information may be limited for a variety of reasons.  Certain brands may have few products in operation resulting in small sample sizes.  Other products may have many observed units in operation but few observed failures or limited time under observation, which makes fitting a parametric lifetime model problematic.  In this paper, we propose the use of Bayesian hierarchical modeling to borrow information across products, thereby improving inferences on lifetime distributions where information is limited by small sample sizes, left-truncation, and right censoring.

The reliability of many engineered products follow a similar pattern. Relatively high rates of failure occurring in early life ("infant mortality") due to defects from the manufacturing process. Following a period of "burn-in", after which the majority of the defective units have failed, failure rates flatten out during a period. Finally, after prolonged use, rates of failure increase due to wearout.  Ignoring this pattern in failure rates can lead to suboptimal decisions and spurious inferences about a product's reliability.   While simple nonparametric methods are available, such as the Kaplan-Meier estimator, these are not suitable for forecasting.  They also do not allow for a straight forward statistical comparisons of different brands in order to select the most reliable product.

@@

While the GLFP model provides flexibility and accords with a macro-level theory of product failure, it requires a lot of data to fit because of its complexity. Bayesian methods offer a means of reducing the amount of data required to fit the model by leveraging multiple sources of information. When multiple populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those with sufficient data. As we will show, hierarchical modeling of the GLFP parameters allows for borrowing of information across populations, imposing "soft" constraints on model parameters. This can enables estimation for all populations; the posterior distribution of time-to-failure for populations with limited data will contain greater uncertainty, while being shrunk toward a "pooled" model. For model fitting, samples obtained from the posterior distribution via Markov Chain Monte Carlo for all model parameters allow easy numerical integration for computing various quantities of interest.\\

@@

\begin{shaded}
  THIS SHOULD BE SEPARATED OUT
  At the other extreme, separately fitting the model to each drive-model, is difficult for drive-models where the data are not strongly informative.  Due to identifiability issues, in these cases the model fit is unstable unless strongly informative priors are chosen.  \\

  We therefore propose to fit the GLFP model using a hierarchical Bayesian approach.  With over 60 drive-models in operation, a hierarchical model is advantageous as allows for borrowing of information to improve estimation of the GLFP parameters for each drive-model.  Moreover, the hierarchical model is a nice compromise between the aggregate and individual modeling approaches: drive-models with lots of failures are estimated precisely; and in cases where little data is available the hierarchical model borrows information from other drive-models to obtain stable estimates.  This allows us to include more of the hard drive data and more importantly, make valid inferences for each group of hard drive brands.\\

  Failure data for series systems may be collected at the system level
  by end-users interested only in the lifetime of the system. \citet{chan} proposed the generalized limited failure population
  model (GLFP), for
  failure times in a population where some units are susceptible to infant mortality. This model
  covers a subset of the class of distributions known as bathtub distributions
  \cite{sujata}. Bathtub distributions are characterized by a U-shaped
  hazard function, where early failures are due to infant mortality and
  late failure are due to wearout. While complex systems actually
  exhibit more than two modes of failure, suitable parametric models
  which can suitably approximate the overall lifetime distribution are
  useful.

  In the case of complex, high-reliability systems where cause of failure is
  not available, a model which
  \begin{enumerate}[a.]
  \item has parameters that can be estimated from the observed data
  \item is flexible enough to fit the observed data
  \item in interpretable with respect to available theory of the underlying process
  \end{enumerate}
  is desired. We present a Bayesian, hierarchical modeling approach for
  grouped failure data. As a benefit of working with Monte Carlo
  samples, we can easily make inference on a wide range of quantities of
  interest while accounting for uncertainty. We demonstrate this approach on real data hard
  drive failure data from a
  cloud-based storage company.
\end{shaded}

To model this mixture of failure distributions in a unified framework, Chan and Meeker proposed the Generalized Limited Failure Population (GLFP) model \citet{chan}.  In addition to describing the failure time distribution for products that are susceptible to infant mortality and wearout failure modes, it theorizes an unknown proportion of defective units which are susceptible to defects.    While the GLFP model provides flexibility over a simple or piece-wise parametric model, it can be difficult to fit with limited data  As Chan and Meeker discuss in their original paper, few observed failures in addition to unknown causes of failure cause maximum likelihood estimates to be unstable or impossible to obtain.  Therefore, when multiple populations are of interest, comparisons based on the standard GLFP model are restricted to those with sufficient data and observed failure from each failure mode.  A naive approach to this limitation is to combine the data from all products and fit a single GLFP model.  This model can be well-estimated, but the pooled GLFP model fits poorly.  At the other extreme, fitting a separate GLFP model to each product is problematic for products where the data are weakly informative.  Due to identifiability issues, in these cases the model fit is unstable unless strongly informative priors are placed on the model parameters.  \\

We therefore propose to fit the GLFP model using a hierarchical Bayesian approach.  A hierarchical model is advantageous as allows for borrowing of information to improve estimation of the GLFP parameters for a whole class of products.  Moreover, the hierarchical model is a nice compromise between the aggregate and individual modeling approaches: product brands with lots of failures are estimated precisely; and in cases where little data is available the hierarchical model borrows information from other products to obtain stable estimates.  This allows us to include all of the data and more importantly, make valid inferences for each individual product.

Failure data for series systems may be collected at the system level
by end-users interested only in the lifetime of the system. \citet{chan} proposed the generalized limited failure population
model (GLFP), for
failure times in a population where some units are susceptible to infant mortality. This model
covers a subset of the class of distributions known as bathtub distributions
\cite{sujata}. Bathtub distributions are characterized by a U-shaped
hazard function, where early failures are due to infant mortality and
late failure are due to wearout. While complex systems actually
exhibit more than two modes of failure, suitable parametric models
which can suitably approximate the overall lifetime distribution are
useful.

In the case of complex, high-reliability systems where cause of failure is
not available, a model which
\begin{enumerate}[a.]
\item has parameters that can be estimated from the observed data
\item is flexible enough to fit the observed data
\item in interpretable with respect to available theory of the underlying process
\end{enumerate}
is desired. We present a Bayesian, hierarchical modeling approach for
grouped failure data. As a benefit of working with Monte Carlo
samples, we can easily make inference on a wide range of quantities of
interest while accounting for uncertainty. We demonstrate this approach on real data hard
drive failure data from a
cloud-based storage company.

@@

\section{Background}
\label{sec:Background}
In engineering applications, a product can often fail from one out of a set of possible malfunctioning components.  For example, a computer system can fail if the mother board, disc drive or power supply stop working.  Circuit boards (CB) can fail due to a manufacturing defect or later as a result of wearout.  The general name for such products is a series system where the lifetime of the product is the minimum failure time across $s$ different components or risks \cite{nelson}.  A common assumption in series systems is the time to failure for each risk, $s$, is statistically independent.  Thus, the overall reliability of a unit is modeled using the product rule across all $s$ risks.  Parameter estimation is straightforward if the cause of failure is known for each observation.  With engineering systems data, however, the exact cause of failure is frequently unknown or masked from the researcher.  \\

Previous papers have employed various data assumptions and methodologies to model masked lifetime failure data.  When modeling computer system failures Reiser et al. assumed each observed failure came from a known subset of failure modes, and estimation was performed using a Bayesian approach \cite{reiser}.  Chan and Meeker labeled the cause of circuit board failures as infant mortality, unknown, or wearout based on the time of observed failures.  This helped identify parameters when using maximum likelihood (ML) estimation.  Extending Chan and Meeker's analysis, Basu et. al performed a Bayesian analysis with informative priors to better identify early versus late failure modes without making any data assumptions \cite{basu}.  Berger and Sun introduced the Poly-Weibull distribution where cause of failure is the minimum of a several Weibull distributions \cite{berger}.  More recently, Ranjan et al. considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions \cite{ranjan}.  Treating the unknown failure modes as incomplete data, an expectation maximization algorithm with ML was used, in addition to Bayesian estimation.

@@

\subsection{Example 1}
We first fit the GLFP model to individual hard drive brand models.  For estimation, we modeled the data on the log scale, using the Smallest Extreme Value (SEV) distribution \cite{meeker}.  If $\text{log}(Y) \sim \text{SEV}(\mu, \sigma)$ then $Y \sim \text{Weibull}(\beta = 1/\sigma, \eta = \exp(\mu))$.  This decision was chosen after encountering numerical issues when performing estimation using the standard Weibull parametrization.  In addition, we transformed the parameters to the lower log quantile, $\log(t_{p_j})$ (where $p_{1} = 0.5$, and $p_2 = 0.2$), and log scale, $\log(\sigma_j)$.  This parameterization was selected as the majority of failures occur at the lower tail of the distribution.  \\

@@

$$t_{p_j} \equiv \mu_{j} + \sigma_{j}\,\Phi^{-1}(p_j) \quad j=1,2$$
\[\log(t_{p_1})  \sim \operatorname{N}(\mu = 7, \sigma = 1 )\]
\[\log(t_{p_2}) \sim \operatorname{N}(\mu = 9, \sigma = 2 )\]
\[\sigma_j \sim \operatorname{Lognormal}(\mu = 0, \sigma =1 ) \quad j = 1,2\]

@@

\section{Data analysis}
\label{sec:Data analysis}
\subsection{Data}
\label{sec:Data}
Backblaze is a company that offers backup storage to protect against data loss.  While selling storage space is Backblaze's main business, since 2013 it has been collecting data on hard drives operating at its facility.  The purpose is to provide consumers and businesses with reliability information on different hard drive-models.  The hard drives continuously spin in controlled storage pods where they run until failure.  When a hard drive fails it is permanently removed, and new hard drives are regularly added to the storage pods.  In addition, the number of storage pods is increasing as Backblaze adds new drive-models to the sample.  Every quarter Backblaze makes its data publicly available through their website \cite{backblaze}. In addition, Backblaze publishes summary statistics of the different hard drive-models currently operating.  No other analysis or modeling of the failure data is provided; however, Backblaze does encourage the public to further analyze its data, which for this paper goes through the first quarter of 2016.

As of the first quarter of 2016, Backblaze was collecting data on 63 different drive-models.  Some drive-models have been running since 2013, while others were added at a later date.  There are a total of 75,297 hard drives in operation.  The distribution of drives by drive-model varies: some drive-models have only have a service record for a single drive whereas the maximum number of service records for a single drive model is 35,860.  Figure \ref{fig1} shows the distribution of total number of failures and total observed running time for drive-models with at least 3 failures.  For model identification, a minimum of 3 failures was the criterion for hard drive brands to be included in the GLFP model.

\begin{figure}[H]
  \includegraphics[width=.9\textwidth]{fig/data-sum.pdf}
  \caption{Left: Total number of failures (log scale) for drive-models with at least 3 observed failures in the Backblaze data. Right: Total observed time in operation (log scale) for each of the drive-models in the Backblaze data.  Both plots are sorted based on the total number of failures.}
  \label{fig1}
\end{figure}

>>>>

\bigskip
\begin{abstract}
Consumers have an interest in accurate assessment of product reliability.  However, available reliability data often contains little information.  Products with long lifetimes produce few observed failures, and when products do fail the cause of failure is often unknown. While computationally convenient, simple parametric models are often limited in modeling this type of data.  Small sample sizes make parameter estimates unstable or difficult to obtain.  Multiple failure modes cause bends in the lifetime distributions that single parametric models cannot capture.

In this paper we present a method for joint estimation of multiple lifetime distributions based on the Generalized Limited Failure Population (GLFP) model \citep{chan}. This 5-parameter model for lifetime data accommodates lifetime distributions with multiple failure modes:  an early failure due to "infant mortality" and a second failure mode due to wearout. We fit the GLFP model using a hierarchical Bayesian approach.  By borrowing strength across populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures.  Moreover, by using a Bayeisan approach, comparison of different product brands is straightforward since the estimation of arbitrary functionals are easily computable from the full posterior distribution.

We illustrate this method on hard drive reliability data from a cloud-based computing company.  The data contains both left-truncation and right-censoring.  Also, the information in the data is unevenly distributed across the populations of interest making a hierarchical approach especially advantageous. 
\end{abstract}

@@

Failure distributions of high-reliability consumer products can be difficult to assess.  Even when field data are available, information may be limited for a variety of reasons.  Certain brands may have few products in operation resulting in small sample sizes.  Other products could have many observed units in operation but few observed failures or limited time under observation, which makes fitting a parametric lifetime model problematic.  In this paper, we propose the use of Bayesian hierarchical modeling to borrow information across products, thereby improving inferences on lifetime distributions where information is limited by small sample sizes, left-truncation, and right censoring.

The reliability of many engineered products follow a similar pattern. Relatively high rates of failure occurring in early life ("infant mortality") due to defects from the manufacturing process. After this "burn-in" period, failure rates stabilize as the majority of defective units have failed and are no longer in the population.  Finally, after prolonged use, rates of failure increase due to wearout.  Ignoring this pattern in failure rates can lead to suboptimal decisions and spurious inferences about a product's reliability.   While simple nonparametric methods are available, such as the Kaplan-Meier estimator, these are not suitable for forecasting.  They also do not allow for a straight forward statistical comparisons of different brands in order to select the most reliable product.

@@

While the GLFP model provides flexibility and accords with a macro-level theory of product failure, it requires a lot of data to fit because of its complexity. Bayesian methods offer a means of reducing the amount of data required to fit the model by leveraging multiple sources of information. When multiple populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those products with sufficient data. As we will show, hierarchical modeling of the GLFP parameters allows for borrowing of information across populations, imposing "soft" constraints on model parameters. This enables estimation for all product groups.  Populations with lots of data will be precisely estimated.  For products with limited data the posterior distribution of time-to-failure will contain greater uncertainty, while being shrunk toward a "pooled" model. Finally, in order to compare the reliability across brands, the Bayesian approach is appealing as samples obtained from the posterior distribution via Markov Chain Monte Carlo allow for easy numerical integration for computing  various quantities of interest. \\

@@

%\begin{shaded}
  %THIS SHOULD BE SEPARATED OUT
  %At the other extreme, separately fitting the model to each drive-model, is difficult for drive-models where the data are not strongly informative.  Due to identifiability issues, in these cases the model fit is unstable unless strongly informative priors are chosen.  \\
  %
  %We therefore propose to fit the GLFP model using a hierarchical Bayesian approach.  With over 60 drive-models in operation, a hierarchical model is advantageous as allows for borrowing of information to improve estimation of the GLFP parameters for each drive-model.  Moreover, the hierarchical model is a nice compromise between the aggregate and individual modeling approaches: drive-models with lots of failures are estimated precisely; and in cases where little data is available the hierarchical model borrows information from other drive-models to obtain stable estimates.  This allows us to include more of the hard drive data and more importantly, make valid inferences for each group of hard drive brands.\\
  %
  %
  %Failure data for series systems may be collected at the system level
  %by end-users interested only in the lifetime of the system. \citet{chan} proposed the generalized limited failure population
  %model (GLFP), for
  %failure times in a population where some units are susceptible to infant mortality. This model
  %covers a subset of the class of distributions known as bathtub distributions
  %\cite{sujata}. Bathtub distributions are characterized by a U-shaped
  %hazard function, where early failures are due to infant mortality and
  %late failure are due to wearout. While complex systems actually
  %exhibit more than two modes of failure, suitable parametric models
  %which can suitably approximate the overall lifetime distribution are
  %useful. 
  %
  %In the case of complex, high-reliability systems where cause of failure is
  %not available, a model which
  %\begin{enumerate}[a.]
  %\item has parameters that can be estimated from the observed data
  %\item is flexible enough to fit the observed data
  %\item in interpretable with respect to available theory of the underlying process
  %\end{enumerate}
  %is desired. We present a Bayesian, hierarchical modeling approach for
  %grouped failure data. As a benefit of working with Monte Carlo
  %samples, we can easily make inference on a wide range of quantities of
  %interest while accounting for uncertainty. We demonstrate this approach on real data hard
  %drive failure data from a
  %cloud-based storage company.
%\end{shaded}
%
%To model this mixture of failure distributions in a unified framework, Chan and Meeker proposed the Generalized Limited Failure Population (GLFP) model \citet{chan}.  In addition to describing the failure time distribution for products that are susceptible to infant mortality and wearout failure modes, it theorizes an unknown proportion of defective units which are susceptible to defects.    While the GLFP model provides flexibility over a simple or piece-wise parametric model, it can be difficult to fit with limited data  As Chan and Meeker discuss in their original paper, few observed failures in addition to unknown causes of failure cause maximum likelihood estimates to be unstable or impossible to obtain.  Therefore, when multiple populations are of interest, comparisons based on the standard GLFP model are restricted to those with sufficient data and observed failure from each failure mode.  A naive approach to this limitation is to combine the data from all products and fit a single GLFP model.  This model can be well-estimated, but the pooled GLFP model fits poorly.  At the other extreme, fitting a separate GLFP model to each product is problematic for products where the data are weakly informative.  Due to identifiability issues, in these cases the model fit is unstable unless strongly informative priors are placed on the model parameters.  \\
%
%We therefore propose to fit the GLFP model using a hierarchical Bayesian approach.  A hierarchical model is advantageous as allows for borrowing of information to improve estimation of the GLFP parameters for a whole class of products.  Moreover, the hierarchical model is a nice compromise between the aggregate and individual modeling approaches: product brands with lots of failures are estimated precisely; and in cases where little data is available the hierarchical model borrows information from other products to obtain stable estimates.  This allows us to include all of the data and more importantly, make valid inferences for each individual product.
%
%
%Failure data for series systems may be collected at the system level
%by end-users interested only in the lifetime of the system. \citet{chan} proposed the generalized limited failure population
%model (GLFP), for
%failure times in a population where some units are susceptible to infant mortality. This model
%covers a subset of the class of distributions known as bathtub distributions
%\cite{sujata}. Bathtub distributions are characterized by a U-shaped
%hazard function, where early failures are due to infant mortality and
%late failure are due to wearout. While complex systems actually
%exhibit more than two modes of failure, suitable parametric models
%which can suitably approximate the overall lifetime distribution are
%useful. 
%
%In the case of complex, high-reliability systems where cause of failure is
%not available, a model which
%\begin{enumerate}[a.]
%\item has parameters that can be estimated from the observed data
%\item is flexible enough to fit the observed data
%\item in interpretable with respect to available theory of the underlying process
%\end{enumerate}
%is desired. We present a Bayesian, hierarchical modeling approach for
%grouped failure data. As a benefit of working with Monte Carlo
%samples, we can easily make inference on a wide range of quantities of
%interest while accounting for uncertainty. We demonstrate this approach on real data hard
%drive failure data from a
%cloud-based storage company.

@@

\section{Background}
\label{sec:Background}
In engineering applications, a product can often fail from one out of a set of possible malfunctioning components.  For example, a computer system can fail if the mother board, disc drive or power supply stop working.  Circuit boards (CB) can fail due to a manufacturing defect or later as a result of wearout.  The general name for such products is a series system where the lifetime of the product is the minimum failure time across $s$ different components or risks \cite{nelson}.  A common assumption in series systems is the time to failure for each risk, $s$, is statistically independent.  Thus, the overall reliability of a unit is modeled using the product rule across all $s$ risks.  Parameter estimation is straightforward if the cause of failure is known for each observation.  With engineering systems data, however, the exact cause of failure is frequently unknown or masked from the analyst.  \\

Previous papers have employed various data assumptions and methodologies to model masked lifetime failure data.  When modeling computer system failures Reiser et al. assumed each observed failure came from a known subset of failure modes, and estimation was performed using a Bayesian approach \cite{reiser}.  Chan and Meeker labeled the cause of circuit board failures as infant mortality, unknown, or wearout based on the time of observed failures.  This helped identify parameters when using maximum likelihood (ML) estimation.  Extending Chan and Meeker's analysis, Basu et.al performed a Bayesian analysis with informative priors to better identify early versus late failure modes without making any data assumptions \cite{basu}.  Berger and Sun introduced the Poly-Weibull distribution where cause of failure is the minimum of a several Weibull distributions \cite{berger}.  More recently, Ranjan et al. considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions \cite{ranjan}.  Treating the unknown failure modes as incomplete data, an expectation maximization (EM) algorithm providing ML estimates was used, in addition to Bayesian estimation.

@@

\subsection{Example 1}
We first demonstrate the GLFP model on real hard drive failure data from a cloud-based storage company.  The data set has over 60 hard drive models in operation, but for now we focus on one model.  For estimation, we modeled the data on the log scale, using the Smallest Extreme Value (SEV) distribution \cite{meeker}.  If $\text{log}(Y) \sim \text{SEV}(\mu, \sigma)$ then $Y \sim \text{Weibull}(\beta = 1/\sigma, \eta = \exp(\mu))$.  This decision was chosen after encountering numerical issues when performing estimation using the standard Weibull parametrization.  In addition, we transformed the parameters to the lower log quantile, $\log(t_{p_j})$ (where $p_{1} = 0.5$, and $p_2 = 0.2$), and log scale, $\log(\sigma_j)$.  This parameterization was selected as the majority of failures occur at the lower tail of the distribution.  \\

@@

$$t_{p_j} \equiv \mu_{j} + \sigma_{j}\,\Phi^{-1}(p_j) \quad j=1,2$$
$$t_{0.5,1} \sim \op{N}(8,4), \quad t_{0.2,2} \sim \op{N}(10,4)$$
$$\mbox{logit}^{-1}\pi_1 \sim \op{Normal}(-3,2),\quad \sigma_1 \sim \op{Lognormal}(0, 2.5), \quad \sigma_2 \sim \op{Lognormal}(0, 2.5) $$

Question: Do we Want to explain these priors here or wait till the hierarchical section where we lay out all the priors?

@@

\section{Data analysis}
\label{sec:Data analysis}
\subsection{Data}
\label{sec:Data}
Backblaze is a company that offers cloud backup storage to protect against data loss.  Since 2013 it has been collecting data on hard drives operating at its facility.  The purpose is to provide consumers and businesses with reliability information on different hard drive-models.  The hard drives continuously spin in controlled storage pods.  Drives are run until failure.  When a hard drive fails it is permanently removed and replaced.  In addition, the number of storage pods is increasing as Backblaze adds to its storage capacity.  Every quarter Backblaze makes its data publicly available through their website \cite{backblaze}. In addition, Backblaze publishes summary statistics of the different hard drive-models currently operating.  No other analysis or modeling of the failure data is provided.  Backblaze does, however, encourage others to further analyze its data.

As of the first quarter of 2016, Backblaze was collecting and reporting data on 63 different drive-models.  Some drive-models have been running since 2013 or before, while others were added at a later date.  Data have been reported on 75,297 different hard drives that are or where in operation.  The distribution of drives by drive-model varies: some drive-models have only have a service record for a single drive whereas the maximum number of daily service records for a single drive model is 35,860.  Figure \ref{fig1} shows the distribution of total number of failures and total observed running time for drive-models with at least 3 failures.  For model identification, a minimum of three failures was the criterion for hard drive brands to be included in our analyses.

\begin{figure}[H]
  \includegraphics[width=.9\textwidth]{fig/data-sum.pdf}
  \caption{Left: Total number of failures (log scale) for drive-models with at least 3 observed failures in the Backblaze data. Right: Total accumulated time in operation (log scale) for each of the drive-models in the Backblaze data.  The plots are sorted based on the total number of failures on the left.}
  \label{fig1}
\end{figure}
